---
title: "simulated_data_final"
output: html_document
date: '2023-02-07'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require('simstudy')
require(ggplot2)
require(EnvStats)
require(truncnorm)
require(psych)
require(ids)
require(data.table)
require(scales)
require(yarrr)
require(vctrs)
library(faux)
library(dplyr)
library(tidyr)
library(faux)
library(bindata)
library(sigmoid)
require(data.table)
#import functions
source('/Users/luciachen/Dropbox/simulated_data/source/R/simulated_data_functions_final.R')
```

## Creating hospital visit count variables
These variables are hospital visit counts, and they are continuous, overlapped and highly correlated. We first created a reference variable using different types of distribution, and then we added random counts to the reference variable so that the succeeding variable is correlated with the preceding variable. We added higher counts to observations with more visits so that the variable in the next time state is highly correlated with the preceding one. Observations with counts > threshold will be added to more counts 

Pareoto variable: https://search.r-project.org/CRAN/refmans/EnvStats/html/

Functions in this script are at simulated_data_functions_final.R

Section 4.2 on the draft
```{r demographic}
#X17: Age variables were indicators: less than 30, 30-39, 40-49, 50-59, 60-69, and above 70.  We generated age using a normal distribution ($m$ = 60, $sd$ = 14.5) with upper and lower bounds (20, 100). Then the normal distribution was converted into groups of indicator variables.

set.seed(100)
num_of_obs = 200000
#We generated age using a normal distribution (m = 60, sd = 14.5) with upper and lower bounds (20, 100) and converted it to groups of indicator variables. 
age <- rtruncnorm(n=num_of_obs, a=20, b=100, mean=60, sd=14.5)

# Then the normal distribution is converted into groups of indicator variables.
agegroup_ls30 <- ifelse(age < 30, 1, 0)
agegroup_ge30 <- ifelse(age >= 30 & age < 40, 1, 0)
agegroup_ge40 <- ifelse(age >= 40 & age < 50, 1, 0)
agegroup_ge50 <- ifelse(age >= 50 & age < 60, 1, 0)
agegroup_ge60 <- ifelse(age >= 60 & age < 70, 1, 0)
agegroup_ge70 <- ifelse(age >= 70, 1, 0)

summary(agegroup_ls30)
sd(agegroup_ls30)
summary(agegroup_ge30)
sd(agegroup_ge30)
summary(agegroup_ge40)
sd(agegroup_ge40)
summary(agegroup_ge50)
sd(agegroup_ge50)
summary(agegroup_ge60)
sd(agegroup_ge60)
summary(agegroup_ge70)
sd(agegroup_ge70)

cor(agegroup_ge30, agegroup_ge60)

#X15: gender [m = 0.0006]
error=rnorm(num_of_obs, 0, 5)
xb <-   -322 + 320*agegroup_ls30  + 320*agegroup_ge30  + 320*agegroup_ge40 + 120*agegroup_ge50 - 560*agegroup_ge60  -  680*agegroup_ge70 +  error
p <- 1/(1 + exp(-xb))

sex_female <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(sex_female)
summary(sex_female)
cor.test(sex_female, agegroup_ls30)
cor.test(sex_female, agegroup_ge30)
cor.test(sex_female, agegroup_ge40)
cor.test(sex_female, agegroup_ge50)
cor.test(sex_female, agegroup_ge60)
cor.test(sex_female, agegroup_ge70)

#X:16 race_white correlated with gender and age    m = 0.71   [-.07, 0.15, -.09]
error=rnorm(num_of_obs, 0, 5)
xb <-   3 - 1*agegroup_ge40 + 2*agegroup_ge70   - 4* sex_female+ error
p <- 1/(1 + exp(-xb))

race_White <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(race_White)
summary(race_White)

cor.test(race_White, agegroup_ge40)
cor.test(race_White, agegroup_ge70)
cor.test(race_White, sex_female)  

```



```{r hospital visits} 
set.seed(100)
num_of_obs = 2000000
#Variable name X1: Count of Emergency Department stops in the last 30 days
Ux_OP_PoC_ED_30days <- get_pareto_var(0.15, 3.4, num_of_obs)   #get_pareto_var parameters:location, shape, num of observations
summary(Ux_OP_PoC_ED_30days)
#add counts to the 30 days variable to create the 730 days variable
# add_counts parameters: 30 days variable, maximum value of the new variable, threshold: values bigger than threshold will have probability of.0000001 to be sampled as none 0, otherwise the probability to be non-zero is .005
Ux_OP_PoC_ED_730days <- add_counts(Ux_OP_PoC_ED_30days, 281, 10) 
#check their correlations
cor(Ux_OP_PoC_ED_30days, Ux_OP_PoC_ED_730days)
# take the difference between the two variables
Ux_OP_PoC_ED <- Ux_OP_PoC_ED_730days - Ux_OP_PoC_ED_30days 

##X2: Count of Urgent Care stops in the last N days
Ux_OP_PoC_UC_30days <- get_pareto_var(0.15, 4, num_of_obs)  #get_pareto_var parameters:location, shape, num
summary(Ux_OP_PoC_UC_30days)
#add counts to the 30 days variable to create the 730 days variable
# add_counts parameters: 30 days variable, maximum value of the new variable, threshold: values bigger than threshold will have probability of.0000001 to be sampled as none 0, otherwise the probability to be non-zero is .005
Ux_OP_PoC_UC_730days <- add_counts(Ux_OP_PoC_UC_30days, 259, 5)
# take the difference between the two variables
Ux_OP_PoC_UC <- Ux_OP_PoC_UC_730days - Ux_OP_PoC_UC_30days

#X3: Count of Tobacco Cessation stops in the last N days, continous, 
Ux_OP_PoC_TobaccoCessation_30day <- get_pareto_var(0.1, 5, num_of_obs)  #get_pareto_var parameters:location, shape, num
summary(Ux_OP_PoC_TobaccoCessation_30day)
#add counts to the 30 days variable to create the 730 days variable, var1: 30 days variable, var2: maximum value of new variable, var3: threshold 
Ux_OP_PoC_TobaccoCessation_730day  <- add_counts(Ux_OP_PoC_TobaccoCessation_30day, 151, 5)
summary(Ux_OP_PoC_TobaccoCessation_730day)
# take the difference between the two variables
Ux_OP_PoC_TobaccoCessation <- Ux_OP_PoC_TobaccoCessation_730day - Ux_OP_PoC_TobaccoCessation_30day
cor.test(Ux_OP_PoC_TobaccoCessation_730day, Ux_OP_PoC_TobaccoCessation_30day)

#X4: Count of Home Based Primary Care (HBPC) stops in the last N days
Ux_OP_MHOC_HBPC_30days <- get_pareto_var(0.01, 3, num_of_obs)   #get_pareto_var parameters:location, shape, num
summary(Ux_OP_MHOC_HBPC_30days)
sd(Ux_OP_MHOC_HBPC_30days)
#add_counts parameters: 30 days variable, maximum value of the new variable, threshold: values bigger than threshold will have probability of.0000001 to be sampled as none 0, otherwise the probability to be non-zero is .005
Ux_OP_MHOC_HBPC_730days  <- add_counts(Ux_OP_MHOC_HBPC_30days, 107, 2)
summary(Ux_OP_MHOC_HBPC_730days )
# take the difference between the two variables
Ux_OP_MHOC_HBPC <- Ux_OP_MHOC_HBPC_730days - Ux_OP_MHOC_HBPC_30days

#: Count of appointments in the last N days var1, pho, threshold, location_var2, shape_var2
#TotalNumOfAppt_91days <- get_correlated_pareto_var(Ux_OP_PoC_ED, 0.20, 1, 4, 3.5)   #parameters:reference var, correlation, threshold (lower than thres -> 0), location, shape, pareoto distribution threshold(lower than thres -> 0, added some noise later on)
TotalNumOfAppt_91days <-get_pareto_var_app(2, 3, num_of_obs) #get_pareto_var parameters:location, shape, num
summary(TotalNumOfAppt_91days)
#add_counts_total parameters: 30 days variable, maximum value of the new variable, threshold: values bigger than threshold will have probability of.000001 {this is the major difference with add_count} to be sampled as none 0, otherwise the probability to be non-zero is .005
TotalNumOfAppt_730days <- add_counts_total(TotalNumOfAppt_91days, 969,50) 
summary(TotalNumOfAppt_730days)
#var1, max, first
## take the difference between the two variables
TotalNumOfAppt <- TotalNumOfAppt_730days - TotalNumOfAppt_91days

```

## binary variables
Here we use functional form to create binary variables

```{r binary, echo=FALSE}
set.seed(100)
#Serious Adverse Event, these events are weakly correlated with hospital visits
#var1, pho, threshold, seed, location_var2, shape_var2, threshold_2

#X_35:Serious Adverse Event -fall
error=rnorm(num_of_obs, 5, 10)
xb <-   -30 + 1.5*Ux_OP_PoC_ED  + 0.04*TotalNumOfAppt + error
p <- 1/(1 + exp(-xb))

Dx_SAE_Falls_0to1_Y   <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Dx_SAE_Falls_0to1_Y )
summary(Dx_SAE_Falls_0to1_Y )
cor.test(Dx_SAE_Falls_0to1_Y, Ux_OP_PoC_ED)
cor.test(Dx_SAE_Falls_0to1_Y, TotalNumOfAppt)

#X_38:Serious Adverse Event - Acetaminophen
error=rnorm(num_of_obs, 2, 10)
xb <-   -75 + 1.5*Ux_OP_PoC_ED_730days  + error
p <- 1/(1 + exp(-xb))

Dx_SAE_Acetaminophen_0to1_Y  <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Dx_SAE_Acetaminophen_0to1_Y  )
cor(Dx_SAE_Acetaminophen_0to1_Y, Ux_OP_PoC_ED_730days)
summary(Dx_SAE_Acetaminophen_0to1_Y )

#X_40:Serious Adverse Event - Poisoning by Sedative
error=rnorm(num_of_obs, 2, 10)
xb <-   -36 + 0.5*Ux_OP_PoC_ED  + error
p <- 1/(1 + exp(-xb))
Dx_SAE_Sedative_0to1_Y  <- rbinom(n = num_of_obs, size = 1, prob = p)
#checking the stats
sum(Dx_SAE_Sedative_0to1_Y)
cor(Dx_SAE_Sedative_0to1_Y, Ux_OP_PoC_ED)
summary(Dx_SAE_Sedative_0to1_Y)


#X_37: Serious Adverse Event - Vehicular Accidents
error=rnorm(num_of_obs, 2, 10)
xb <-   -28 + 0.8*Ux_OP_PoC_ED  + error
p <- 1/(1 + exp(-xb))
Dx_SAE_Vehicle_0to1_Y  <- rbinom(n = num_of_obs, size = 1, prob = p)
#checking the stats
sum(Dx_SAE_Vehicle_0to1_Y)
cor(Dx_SAE_Vehicle_0to1_Y, Ux_OP_PoC_ED_730days)
summary(Dx_SAE_Vehicle_0to1_Y)

#X_36:Serious Adverse Event - Other Accidents, no correlation with other vars m = 0.002
error=rnorm(num_of_obs, 2, 10)
xb <-   -30 + 1.5*Ux_OP_PoC_ED  + error
p <- 1/(1 + exp(-xb))
Dx_SAE_OtherAccident_0to1_Y  <- rbinom(n = num_of_obs, size = 1, prob = p)
#checking the stats
summary(Dx_SAE_OtherAccident_0to1_Y)
sum(Dx_SAE_OtherAccident_0to1_Y)
cor(Dx_SAE_OtherAccident_0to1_Y, Ux_OP_PoC_ED)


#X_39:Serious Adverse Event - Poisoning by Other drugs  m = 0.0008
error=rnorm(num_of_obs, 2, 10)
xb <-   -36 + 1*Ux_OP_PoC_ED  + error
p <- 1/(1 + exp(-xb))
Dx_SAE_OtherDrug_0to1_Y  <- rbinom(n = num_of_obs, size = 1, prob = p)
#checking the stats
summary(Dx_SAE_OtherDrug_0to1_Y)
sum(Dx_SAE_OtherDrug_0to1_Y)
cor(Dx_SAE_OtherDrug_0to1_Y, Ux_OP_PoC_ED)

#X6: Insomnia - Patient on drug in the past 365 days
#Insomnia medications prescribed in the past 365 days, this is a random variable with binomial distribution with a probability of success of 0.0004 on each trial.
Rx_Insomnia_0to365_D <-  rbinom(num_of_obs, 1,.0004)
summary(Rx_Insomnia_0to365_D)

#X7: Insomnia - Patient on drug at time of selection date
error=rnorm(num_of_obs, 2, 10)
xb <-   -30 + 95*Rx_Insomnia_0to365_D  + error
p <- 1/(1 + exp(-xb))

Paton_Rx_Insomnia  <- rbinom(n = num_of_obs, size = 1, prob = p)
#checking the stats
summary(Paton_Rx_Insomnia)
cor(Paton_Rx_Insomnia, Rx_Insomnia_0to365_D)

#X8: methadone prescription, long/short act, no corr with other vars
Rx_OAT_methadone_12m <-rbinom(num_of_obs, 1,.0001)
summary(Rx_OAT_methadone_12m)

#X9: buprenorphine prescription
Rx_OAT_buprenorphine_12m <-rbinom(num_of_obs, 1,.0006)
summary(Rx_OAT_buprenorphine_12m)

#X14: Alcohol Use Disorder [0.14, 0.15, 19]   m = [0.05]
error=rnorm(num_of_obs, 2, 10)
xb <-   -28 + 1.8*Ux_OP_PoC_ED  +  1.2*TotalNumOfAppt + error
p <- 1/(1 + exp(-xb))

Dx_AlcoholUD_0to1_Y  <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Dx_AlcoholUD_0to1_Y )
summary(Dx_AlcoholUD_0to1_Y)
cor(Dx_AlcoholUD_0to1_Y, Ux_OP_PoC_ED )
#cor(Dx_AlcoholUD_0to1_Y, MOPROB_AllAppt)
cor(Dx_AlcoholUD_0to1_Y, TotalNumOfAppt)


#X11: Substance use disorders not included in other definitions in the dataset. [0.23, 0.14, 0.14] m = [0.008]
error=rnorm(num_of_obs, 2, 10)
xb <-   -28 +  2.5*Dx_AlcoholUD_0to1_Y  + 1.5*Ux_OP_PoC_ED    + 1*TotalNumOfAppt  + error
p <- 1/(1 + exp(-xb))

Dx_SUD_CatchAll_RV2_0to1_Y  <- rbinom(n = num_of_obs, size = 1, prob = p)
#checking the stats
sum(Dx_SUD_CatchAll_RV2_0to1_Y)
summary(Dx_SUD_CatchAll_RV2_0to1_Y)
cor(Dx_AlcoholUD_0to1_Y, Dx_SUD_CatchAll_RV2_0to1_Y)
cor(Ux_OP_PoC_ED, Dx_SUD_CatchAll_RV2_0to1_Y)
cor(TotalNumOfAppt, Dx_SUD_CatchAll_RV2_0to1_Y)


#X12: Sedative Use Disorder [0.15, 0.1, 0.08] m = 0.002
error=rnorm(num_of_obs, 2, 10)
xb <-   -32 + 10*Dx_SUD_CatchAll_RV2_0to1_Y + 8*Dx_AlcoholUD_0to1_Y + 0.6*Ux_OP_PoC_ED + error
p <- 1/(1 + exp(-xb))

Dx_SedativeUD_0to1_Y    <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Dx_SedativeUD_0to1_Y)
summary(Dx_SedativeUD_0to1_Y  )
cor(Dx_SedativeUD_0to1_Y, Dx_SUD_CatchAll_RV2_0to1_Y)
cor(Dx_SedativeUD_0to1_Y, Dx_AlcoholUD_0to1_Y)
cor(Dx_SedativeUD_0to1_Y,Ux_OP_PoC_ED)


#X13: Amphetamine and other Stimulant Disorder  [0.15, 0.1, 0.1, 0.1, 0.28] m = 0.005
error=rnorm(num_of_obs, 2, 10) 
xb <-   -32 + 10.5*Dx_SedativeUD_0to1_Y + 1*Ux_OP_PoC_ED   + 0.2*TotalNumOfAppt + 10.5*Dx_SUD_CatchAll_RV2_0to1_Y + error
p <- 1/(1 + exp(-xb))

Dx_Amphetamine_OtherStim_UD_0to1  <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Dx_Amphetamine_OtherStim_UD_0to1)
summary(Dx_Amphetamine_OtherStim_UD_0to1)

cor.test(Dx_Amphetamine_OtherStim_UD_0to1, Dx_SedativeUD_0to1_Y )
cor.test(Dx_Amphetamine_OtherStim_UD_0to1, Ux_OP_PoC_ED_730days)
cor.test(Dx_Amphetamine_OtherStim_UD_0to1, TotalNumOfAppt_730days)
cor.test(Dx_Amphetamine_OtherStim_UD_0to1, Dx_SUD_CatchAll_RV2_0to1_Y)


```


#variables from the STORM dataset
```{r storm, echo=FALSE}
set.seed(100)
#X41-52 MME  Morphine equivalent daily dose 10 months ago, continous ,  these variables are accumulative 
c<- rbeta(num_of_obs,1,150) #generate beta distribution
c2 <- rescale(c, to = c(0, 5000)) #rescale the beta distribution
Rx_Medd <- add_random_zero(c2) #add 0 randomly to beta distribution
Rx_Medd_1m <- add_Medd(Rx_Medd, probs=  c(0.9999, 0.0001), 5000) # parameters: probs of adding 0, prob of adding a random number, bound of the random number
Rx_Medd_2m <- add_Medd(Rx_Medd_1m, probs=  c(0.9999, 0.0001), 8000)
Rx_Medd_3m <- add_Medd(Rx_Medd_2m, probs=  c(0.9999, 0.0001), 8000)
Rx_Medd_4m <- add_Medd(Rx_Medd_3m, probs=  c(0.9999, 0.0001), 5000)
Rx_Medd_5m <- add_Medd(Rx_Medd_4m, probs=  c(0.9999, 0.0001), 5000)
Rx_Medd_6m <- add_Medd(Rx_Medd_5m, probs=  c(0.9999, 0.0001), 5000)
Rx_Medd_7m <- add_Medd(Rx_Medd_6m, probs=  c(0.9999, 0.0001), 5000)
Rx_Medd_8m <- add_Medd(Rx_Medd_7m, probs=  c(0.9999, 0.0001), 8000)
Rx_Medd_9m <- add_Medd(Rx_Medd_8m, probs=  c(0.9999, 0.0001), 8000)
Rx_Medd_10m <- add_Medd(Rx_Medd_9m, probs=  c(0.9999, 0.0001), 8000)
Rx_Medd_11m <- add_Medd(Rx_Medd_10m, probs=  c(0.9999, 0.0001), 8000)


#X18: Urine drug screen for heroin/morphine past year
#correlated but not overlapping
Dx_UDS1_12m <- rbinom(num_of_obs, 1,.08)
#X19: Urine drug screen for nonmorphine opioid compounds past year
error=rnorm(num_of_obs, 2, 10)
xb <-   -30 + 35*Dx_UDS1_12m + error
p <- 1/(1 + exp(-xb))

Dx_UDS2_12m    <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Dx_UDS2_12m)
summary(Dx_UDS2_12m)
cor(Dx_UDS2_12m, Dx_UDS1_12m)


#X20: Urine drug screen for nonopioid abusable substances past year
error=rnorm(num_of_obs, 2, 10)
xb <-   -22 + 35*Dx_UDS1_12m + error
p <- 1/(1 + exp(-xb))

Dx_UDS3_12m <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Dx_UDS3_12m)
summary(Dx_UDS3_12m)
cor(Dx_UDS3_12m, Dx_UDS1_12m)

#X21: care at pain clinic, mildly related to Medd 0.13, 0.21 correlated with Dx_USD, binary
error=rnorm(num_of_obs, 0, 3)
xb <-   -55 + 1*Rx_Medd + 1*Rx_Medd_10m + 1*Rx_Medd_11m + 1*Rx_Medd_1m + 1*Rx_Medd_2m + 1*Rx_Medd_3m + 1*Rx_Medd_4m + 1*Rx_Medd_5m + 1*Rx_Medd_6m + 1*Rx_Medd_7m + 1*Rx_Medd_8m + 1*Rx_Medd_9m + 18*Dx_UDS1_12m +18* Dx_UDS1_12m + 18*Dx_UDS1_12m + error
p <- 1/(1 + exp(-xb))

Ux_RehMedPain  <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Ux_RehMedPain)
summary(Ux_RehMedPain)
cor.test(Ux_RehMedPain, Rx_Medd)
cor.test(Ux_RehMedPain, Dx_UDS1_12m)


#X28: Opioids For Pains in the past 0 to 182 days (binary) 
error=rnorm(num_of_obs, 10, 5)
xb <-  -15 + 0.8* Ux_OP_PoC_ED + error
p <- 1/(1 + exp(-xb))

Rx_OpioidForPain_0to182_D <- rbinom(n = num_of_obs, size = 1, prob = p)

#print statistics
sum(Rx_OpioidForPain_0to182_D)
sd(Rx_OpioidForPain_0to182_D)
summary(Rx_OpioidForPain_0to182_D)
cor.test(Rx_OpioidForPain_0to182_D, Ux_OP_PoC_ED_730days)


#X29 Opioids For Pains in the past 182 to 365 days binary  cor[0.62, 0.21] m, sd = [0.12, 0.33]
xb <-  -17 + 0.4*Rx_OpioidForPain_0to182_D  +  0.8* Ux_OP_PoC_ED  + error
p <- 1/(1 + exp(-xb))

Rx_OpioidForPain_182to365_D <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Rx_OpioidForPain_182to365_D)
summary(Rx_OpioidForPain_182to365_D)
cor.test(Rx_OpioidForPain_182to365_D, Rx_OpioidForPain_0to182_D)
cor.test(Rx_OpioidForPain_182to365_D, Ux_OP_PoC_ED_730days)



#pain conditions
#X25: Urogenital pain
#not related to anything
Dx_Pain_Urogenital_0to1_Y <- rbinom(num_of_obs, 1,.014)

# Dx_Pain_Neuropathy_0to1_Y mildly correlated with Rx_OpioidForPain_0to182_D 0.11, Paton_Rx_OpioidForPain 0.08, age 40-70 about 0.12     m =.078
#X22: neuropathy pain
error=rnorm(num_of_obs, 0, 3)
xb <-   -12  + 1*Rx_OpioidForPain_0to182_D + 1*Rx_OpioidForPain_182to365_D + 7*agegroup_ge50 + 7*agegroup_ge60  +  7*agegroup_ge70  + error
p <- 1/(1 + exp(-xb))

Dx_Pain_Neuropathy_0to1_Y  <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Dx_Pain_Neuropathy_0to1_Y)
summary(Dx_Pain_Neuropathy_0to1_Y)
cor.test(Dx_Pain_Neuropathy_0to1_Y, Rx_OpioidForPain_182to365_D)
cor.test(Dx_Pain_Neuropathy_0to1_Y, agegroup_ge50)

#X23 back pain
#midly correlated with TotalNumOfAppt 0.13- 0.18, Paton_Rx_OpioidForPain, Rx_OpioidForPain_0to182_D 0.22  m =0.22
error=rnorm(num_of_obs, 8, 8)
xb <-   -20  + 4*Rx_OpioidForPain_0to182_D + 3*Rx_OpioidForPain_182to365_D +  0.5*TotalNumOfAppt_730days + error
p <- 1/(1 + exp(-xb))

Dx_Pain_Back <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Dx_Pain_Back)
summary(Dx_Pain_Back)
cor.test(Dx_Pain_Back, Rx_OpioidForPain_182to365_D)
cor.test(Dx_Pain_Back, TotalNumOfAppt)

#X24 neck pain
#Ux_OP_PoC_ED_730days 0.1, Rx_OpioidForPain_0to182_D 0.14, Paton_Rx_OpioidForPain 0.10, TotalNumOfAppt_91days 0.14- 0.17   m = 0.06
error=rnorm(num_of_obs, -2, 8)
xb <-   -19 + 2*Ux_OP_PoC_UC  + 0.5*TotalNumOfAppt   + 2.8*Rx_OpioidForPain_0to182_D + 1.8*Rx_OpioidForPain_182to365_D + error
p <- 1/(1 + exp(-xb))

Dx_Pain_Neck_0to1_Y <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Dx_Pain_Neck_0to1_Y )
summary(Dx_Pain_Neck_0to1_Y)
cor.test(Dx_Pain_Neck_0to1_Y, Ux_OP_PoC_UC)
cor.test(Dx_Pain_Neck_0to1_Y, TotalNumOfAppt)
cor.test(Dx_Pain_Neck_0to1_Y, Rx_OpioidForPain_0to182_D)

#X26 Fibromyalgia pain
error=rnorm(num_of_obs, 0, 5)
xb <-   -12 + 0.05*TotalNumOfAppt + error
p <- 1/(1 + exp(-xb))


Dx_Pain_Fibromyalgia_0to1_Y     <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Dx_Pain_Fibromyalgia_0to1_Y)
summary(Dx_Pain_Fibromyalgia_0to1_Y)
cor.test(Dx_Pain_Fibromyalgia_0to1_Y, TotalNumOfAppt)

#X27 Endometriosis no corr with other var
Dx_Endometriosis_0to1_Y <- rbinom(num_of_obs, 1,.0006)
summary(Dx_Endometriosis_0to1_Y)


# mental illnesses
#X30 psychoses past year, this one is related to age (-0.14) and substance use such as alcohol(0.11) cannabis, cocaine, nicotine, depression
error=rnorm(num_of_obs, 0, 2)
xb <-   -1 + 8*Dx_AlcoholUD_0to1_Y  - 18*agegroup_ge50 - 20*agegroup_ge60  -  45*agegroup_ge70 + error
p <- 1/(1 + exp(-xb))

EH_Psychoses <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(EH_Psychoses)
summary(EH_Psychoses)
cor.test(EH_Psychoses, Dx_AlcoholUD_0to1_Y)
cor.test(EH_Psychoses, agegroup_ge70)

#X31 depression, correlated with alcoho, psychosis, homelessness, elixhausr drug abuse
error=rnorm(num_of_obs, 0, 5)
xb <-   -20 + 2*Dx_AlcoholUD_0to1_Y  + 27*EH_Psychoses + error
p <- 1/(1 + exp(-xb))

Dx_MDD_12m  <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Dx_MDD_12m)
summary(Dx_MDD_12m )
cor.test(Dx_MDD_12m, Dx_AlcoholUD_0to1_Y)
cor.test(Dx_MDD_12m, EH_Psychoses)



#X32 PTSD m = 0.12
error=rnorm(num_of_obs, 0, 5)
xb <-   -15 + 1*Ux_OP_PoC_ED + 0.6*TotalNumOfAppt + error
p <- 1/(1 + exp(-xb))

Dx_PTSD_1to2_Y  <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Dx_PTSD_1to2_Y)
summary(Dx_PTSD_1to2_Y)
cor.test(Dx_PTSD_1to2_Y, Ux_OP_PoC_ED)
cor.test(Dx_PTSD_1to2_Y, TotalNumOfAppt)



#X10: anxiolytic [0.14, 0.20, 0.22] m = 0.1
error=rnorm(num_of_obs, 0, 5) 
xb <-   -12 + 0.7*Ux_OP_PoC_ED + 0.5*TotalNumOfAppt + 1.2*Dx_PTSD_1to2_Y + error
p <- 1/(1 + exp(-xb))

Rx_Anxiolytic_0to365_D   <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Rx_Anxiolytic_0to365_D )
summary(Rx_Anxiolytic_0to365_D )
cor.test(Rx_Anxiolytic_0to365_D , Ux_OP_PoC_ED)
cor.test(Rx_Anxiolytic_0to365_D , TotalNumOfAppt)
cor.test(Rx_Anxiolytic_0to365_D , Dx_PTSD_1to2_Y)


#X33 homelessness [0.16, 0.10, 0.16, 0.15, 15] m = 0.02
error=rnorm(num_of_obs, 0, 5)
xb <-   -7 + 0.2*Ux_OP_PoC_ED + 0.005*TotalNumOfAppt + 1.5*Dx_SUD_CatchAll_RV2_0to1_Y + 1*Dx_AlcoholUD_0to1_Y + Dx_Amphetamine_OtherStim_UD_0to1
p <- 1/(1 + exp(-xb))
 

Dx_Homeless_1to2_Y <-  rbinom(n = num_of_obs, size = 1, prob = p)
sum(Dx_Homeless_1to2_Y)
summary(Dx_Homeless_1to2_Y)
cor.test(Dx_Homeless_1to2_Y, Ux_OP_PoC_ED)
cor.test(Dx_Homeless_1to2_Y, TotalNumOfAppt) 
cor.test(Dx_Homeless_1to2_Y, Dx_SUD_CatchAll_RV2_0to1_Y) 
cor.test(Dx_Homeless_1to2_Y, Dx_AlcoholUD_0to1_Y) 
cor.test(Dx_Homeless_1to2_Y, Dx_Amphetamine_OtherStim_UD_0to1) 


#X34: obesity  m = 0.13  [0.17]
error=rnorm(num_of_obs, 0, 5)
xb <-   -9 + 0.3*TotalNumOfAppt + error
p <- 1/(1 + exp(-xb))

Dx_EH_Obesity_1to2_Y <-   rbinom(n = num_of_obs, size = 1, prob = p)
sum(Dx_EH_Obesity_1to2_Y)
summary(Dx_EH_Obesity_1to2_Y)
cor.test(Dx_EH_Obesity_1to2_Y, TotalNumOfAppt) 

```


```{r outcome}
# ## build Y variable
error=rnorm(num_of_obs, 0, 50)
xb <- -380 +  1*Ux_OP_PoC_ED + 10*Ux_OP_PoC_UC + 30*Ux_OP_PoC_TobaccoCessation  + 3*Ux_OP_MHOC_HBPC + 3.5*TotalNumOfAppt  +   3*Rx_Insomnia_0to365_D + 35*Paton_Rx_Insomnia + 2*Rx_OAT_methadone_12m + 2*Rx_OAT_buprenorphine_12m + 0.5*Rx_Anxiolytic_0to365_D + 1*Dx_SUD_CatchAll_RV2_0to1_Y + 3*Dx_SedativeUD_0to1_Y + 1*Dx_Amphetamine_OtherStim_UD_0to1 +  1*Dx_AlcoholUD_0to1_Y + 0.004*agegroup_ls30 + 28*agegroup_ge30 + 28*agegroup_ge40 + 28*agegroup_ge50 + 28*agegroup_ge60 - 30*agegroup_ge70 - 25*sex_female + 43*race_White + 20*Dx_UDS1_12m + 20*Dx_UDS2_12m+ 20*Dx_UDS3_12m + 20*Ux_RehMedPain +35*Dx_Pain_Neuropathy_0to1_Y  + 35*Dx_Pain_Back + 35*Dx_Pain_Neck_0to1_Y + 1*Dx_Pain_Urogenital_0to1_Y + 1*Dx_Pain_Fibromyalgia_0to1_Y+ 20*Dx_Endometriosis_0to1_Y + 5*Rx_OpioidForPain_0to182_D + 10*Rx_OpioidForPain_182to365_D + 5*EH_Psychoses  + 20*Dx_MDD_12m + 0.5*Dx_PTSD_1to2_Y + 0.005*Dx_Homeless_1to2_Y +  0.5*Dx_EH_Obesity_1to2_Y + 20*Dx_SAE_Falls_0to1_Y + 20*Dx_SAE_OtherAccident_0to1_Y + 60*Dx_SAE_Vehicle_0to1_Y + 3*Dx_SAE_Acetaminophen_0to1_Y + 30*Dx_SAE_OtherDrug_0to1_Y + 30*Dx_SAE_Sedative_0to1_Y + 0.005*Rx_Medd + 0.001*Rx_Medd_1m +0.001*Rx_Medd_2m + 0.001*Rx_Medd_3m + 0.001*Rx_Medd_4m + 0.001*Rx_Medd_5m + 0.003*Rx_Medd_6m + 0.003*Rx_Medd_7m + 0.005*Rx_Medd_8m + 0.004*Rx_Medd_9m + 0.005*Rx_Medd_10m + 0.003*Rx_Medd_11m + error


p <- 1/(1 + exp(-xb))
summary(p)
 
Dx_OpioidOverdose_0to1_Y <- rbinom(n = num_of_obs, size = 1, prob = p)
sum(Dx_OpioidOverdose_0to1_Y)
summary(Dx_OpioidOverdose_0to1_Y)
#checking the correlations
# cor.test(Dx_OpioidOverdose_0to1_Y, Dx_Pain_Fibromyalgia_0to1_Y)
# cor.test(Dx_OpioidOverdose_0to1_Y, Dx_Endometriosis_0to1_Y)
# cor.test(Dx_OpioidOverdose_0to1_Y, Dx_EH_Obesity_1to2_Y)
# cor.test(Dx_OpioidOverdose_0to1_Y, Dx_Homeless_1to2_Y)
# cor.test(Dx_OpioidOverdose_0to1_Y, Dx_PTSD_1to2_Y)
# cor.test(Dx_OpioidOverdose_0to1_Y, Rx_OAT_methadone_12m)
# cor.test(Dx_OpioidOverdose_0to1_Y,  Rx_OAT_buprenorphine_12m)
# cor.test(Dx_OpioidOverdose_0to1_Y,  Rx_Anxiolytic_0to365_D)


# #combine all variables to generate the simulated dataset, list x number in here, in the paper we define the one-hot encoding variables as one variable
simulated_data <- cbind(Ux_OP_PoC_ED, Ux_OP_PoC_UC,  Ux_OP_PoC_TobaccoCessation,  TotalNumOfAppt, Ux_OP_MHOC_HBPC, Rx_OpioidForPain_0to182_D, Rx_OpioidForPain_182to365_D,  Rx_Insomnia_0to365_D, Paton_Rx_Insomnia, Dx_SAE_Acetaminophen_0to1_Y, Dx_SAE_Falls_0to1_Y, Dx_SAE_OtherAccident_0to1_Y, Dx_SAE_OtherDrug_0to1_Y, Dx_SAE_Sedative_0to1_Y, Dx_SAE_Vehicle_0to1_Y, Dx_SedativeUD_0to1_Y, Dx_SUD_CatchAll_RV2_0to1_Y, Dx_AlcoholUD_0to1_Y, Dx_Amphetamine_OtherStim_UD_0to1, Rx_Medd, Rx_Medd_10m, Rx_Medd_11m, Rx_Medd_1m, Rx_Medd_2m, Rx_Medd_3m, Rx_Medd_4m, Rx_Medd_5m, Rx_Medd_6m, Rx_Medd_7m, Rx_Medd_8m,Rx_Medd_9m, agegroup_ls30, agegroup_ge30, agegroup_ge40, agegroup_ge50, agegroup_ge60, agegroup_ge70, sex_female, race_White,  Dx_UDS1_12m, Dx_UDS2_12m, Dx_UDS3_12m,  EH_Psychoses, Dx_MDD_12m, Dx_Pain_Urogenital_0to1_Y, Dx_Pain_Neuropathy_0to1_Y, Dx_Pain_Back, Dx_Pain_Neck_0to1_Y, Dx_Pain_Fibromyalgia_0to1_Y, Dx_Endometriosis_0to1_Y, Dx_PTSD_1to2_Y,Dx_Homeless_1to2_Y, Dx_EH_Obesity_1to2_Y, Rx_Anxiolytic_0to365_D,  Rx_OAT_methadone_12m, Rx_OAT_buprenorphine_12m, Dx_OpioidOverdose_0to1_Y)


#generate stats and correlation table, we want a different order of these variables in the table to match with the variable orders in the stats table
simulated_data_for_stats <- cbind(Ux_OP_PoC_ED_30days, Ux_OP_PoC_ED_730days, Ux_OP_PoC_UC_30days, Ux_OP_PoC_UC_730days, Ux_OP_PoC_TobaccoCessation_30day,  Ux_OP_PoC_TobaccoCessation_730day, TotalNumOfAppt_91days, TotalNumOfAppt_730days,  Ux_OP_MHOC_HBPC_30days, Ux_OP_MHOC_HBPC_730days, Rx_OpioidForPain_0to182_D, Rx_OpioidForPain_182to365_D,  Rx_Insomnia_0to365_D, Paton_Rx_Insomnia, Dx_SAE_Acetaminophen_0to1_Y, Dx_SAE_Falls_0to1_Y, Dx_SAE_OtherAccident_0to1_Y, Dx_SAE_OtherDrug_0to1_Y, Dx_SAE_Sedative_0to1_Y, Dx_SAE_Vehicle_0to1_Y, Dx_SedativeUD_0to1_Y, Dx_SUD_CatchAll_RV2_0to1_Y, Dx_AlcoholUD_0to1_Y, Dx_Amphetamine_OtherStim_UD_0to1, Rx_Medd, Rx_Medd_10m, Rx_Medd_11m, Rx_Medd_1m, Rx_Medd_2m, Rx_Medd_3m, Rx_Medd_4m, Rx_Medd_5m, Rx_Medd_6m, Rx_Medd_7m, Rx_Medd_8m,Rx_Medd_9m, agegroup_ls30, agegroup_ge30, agegroup_ge40, agegroup_ge50, agegroup_ge60, agegroup_ge70, sex_female, race_White,  Dx_UDS1_12m, Dx_UDS2_12m, Dx_UDS3_12m, EH_Psychoses, Dx_MDD_12m, Dx_PTSD_1to2_Y, Dx_Homeless_1to2_Y, Dx_EH_Obesity_1to2_Y, Dx_Pain_Urogenital_0to1_Y, Dx_Pain_Neuropathy_0to1_Y, Dx_Pain_Back, Dx_Pain_Neck_0to1_Y, Dx_Pain_Fibromyalgia_0to1_Y, Dx_Endometriosis_0to1_Y, Rx_Anxiolytic_0to365_D,  Rx_OAT_methadone_12m, Rx_OAT_buprenorphine_12m, Dx_OpioidOverdose_0to1_Y)

```


```{r write files}
#write.csv(simulated_data, "/Users/luciachen/Desktop/simulated_data/simulated_data_big_sample.csv")
#convert variables to dataframe and data table
simulated_data <- as.data.frame(simulated_data)
simulated_data <- data.table(simulated_data)

#change variale names, here we change the pain conditions and mental illenss with a different prefix for variable selection in the ML pipeline 
#EH_Psychoses, Dx_MDD_12m, Dx_PTSD_1to2_Y, Dx_Homeless_1to2_Y, Dx_EH_Obesity_1to2_Y, Dx_Pain_Urogenital_0to1_Y, Dx_Pain_Neuropathy_0to1_Y, Dx_Pain_Back, Dx_Pain_Neck_0to1_Y, Dx_Pain_Fibromyalgia_0to1_Y, Dx_Endometriosis_0to1_Y,
simulated_data  <- simulated_data %>% rename("MH_Psychoses" = "EH_Psychoses",
                                        "MH_MDD_12m" = "Dx_MDD_12m",
                                        "MH_PTSD_1to2_Y" = "Dx_PTSD_1to2_Y", 
                                        "EH_Obesity_1to2_Y" = "Dx_EH_Obesity_1to2_Y",
                                        "EH_Homeless_1to2_Y" = "Dx_Homeless_1to2_Y",
                                        "PA_Pain_Urogenital_0to1_Y" = "Dx_Pain_Urogenital_0to1_Y",
                                        "PA_Pain_Neuropathy_0to1_Y" = "Dx_Pain_Neuropathy_0to1_Y",
                                        "PA_Pain_Back" = "Dx_Pain_Back",
                                        "PA_Pain_Neck_0to1_Y" = "Dx_Pain_Neck_0to1_Y",
                                        "PA_Pain_Fibromyalgia_0to1_Y" = "Dx_Pain_Fibromyalgia_0to1_Y",
                                        "PA_Endometriosis_0to1_Y" = "Dx_Endometriosis_0to1_Y")

#convert the stats variables into tables
simulated_data_for_stats <- as.data.frame(simulated_data_for_stats)
simulated_data_for_stats <- data.table(simulated_data_for_stats)

#get correlations between variables 
simulated_data_corr <- round(cor(simulated_data_for_stats), digits = 4)

#get data statistics
simulated_stats <- describe(simulated_data_for_stats, quant=c(.1, .2, .3, .4, .5, .6, .7, .8, .9), skew = FALSE)
simulated_stats <- within(simulated_stats, rm("se", "range", "vars", "n"))

# transpose metrics
t_simulated_stats<- t(simulated_stats)

# get row and colnames in order
colnames(t_simulated_stats) <- rownames(simulated_stats)
rownames(t_simulated_stats) <- colnames(simulated_stats)

#store all the stats results
write.csv(simulated_data, "/Users/luciachen/Desktop/Desktop_MacbookPro/simulated_data/simulated_data_big_sample.csv") #data for ML pipeline
write.csv(simulated_data_corr, "/Users/luciachen/Desktop/Desktop_MacbookPro/simulated_data/simulated_data_corr.csv") #correlation
write.csv(t_simulated_stats, "/Users/luciachen/Desktop/Desktop_MacbookPro/simulated_data/simulated_stats.csv") #stats 


#select a subset of correlation from the table
corr_matrix <- read.csv("/Users/luciachen/Desktop/Desktop_MacbookPro/simulated_data/simulated_data_corr.csv")
corr_matrix <- data.frame(corr_matrix, row.names = 1)


#In this section we create a matrix for the storm variables to study the correlations between these variables.
#storm vars
var_names_storm <- c("Dx_UDS1_12m", "Dx_UDS2_12m", "Dx_UDS3_12m", "Rx_Medd", "Rx_Medd_10m", "Rx_Medd_11m", "Rx_Medd_1m", "Rx_Medd_2m", "Rx_Medd_3m", "Rx_Medd_4m", "Rx_Medd_5m", "Rx_Medd_6m", "Rx_Medd_7m", "Rx_Medd_8m", "Rx_Medd_9m","Dx_MDD_12m", "EH_Arrhythmia_12m", "EH_BlAnemia_12m", "EH_ChrnPulm_12m", "EH_Psychoses","EH_RheumArt_12m")

#var_names_storm <- c("Dx_UDS1_12m", "Dx_UDS2_12m", "Dx_UDS3_12m", "Rx_Medd", "Rx_Medd_10m", "Rx_Medd_11m", "Rx_Medd_1m", "Rx_Medd_2m", "Rx_Medd_3m", "Rx_Medd_4m", "Rx_Medd_5m", "Rx_Medd_6m", "Rx_Medd_7m", "Rx_Medd_8m", "Rx_Medd_9m","Dx_MDD_12m", "EH_Arrhythmia_12m", "EH_BlAnemia_12m", "EH_ChrnPulm_12m", "EH_Psychoses_12m", "EH_RheumArt_12m")


selected_cor_storm <- corr_matrix[var_names_storm,var_names_storm]


#corr_max_storm <- cov2cor(as.matrix(covariance_max_storm))
write.csv(selected_cor_storm, "/Users/luciachen/Dropbox/simulated_data/sim_corr_max_storm_selected.csv")

```






Task 2: shift data distribution 
```{r datashift}
#attempt 1: here we reduce the mean of the variable by certain percentage
reduction <- function (val, redu) {
  #this function reduce a value by certain amount
  if(val > 0) {
      new_val <- val - redu
  }
  else{
    new_val = 0
  }
  return (new_val)
}

transform_mean <- function(percent, var){
  #number of observation with y > 0
  N1 <- sum(var != 0) 
  #new value = mean * percent that we want to reduce, create a vector to be deducted by the original variable. 
  reduce  <- mean(var)*percent
  print(reduce)
  #normalize the vector
  reduce_scaled <- reduce*(length(var)/N1)
  print(reduce_scaled)
  #apply reduction function on original variable, t
  var_new <- unlist(mapply(FUN=reduction, val=var, redu = reduce_scaled, USE.NAMES=FALSE))
  
    
  return (var_new)
}

reduce_precentage = 0.1 #suppose we reduce the mean by 3 percent
Rx_Medd_new <- transform_mean(reduce_precentage, Rx_Medd)
Rx_Medd_10m_new <- transform_mean(reduce_precentage, Rx_Medd_10m)
Rx_Medd_11m_new <-transform_mean(reduce_precentage, Rx_Medd_11m)
Rx_Medd_1m_new <- transform_mean(reduce_precentage, Rx_Medd_10m)
Rx_Medd_2m_new <- transform_mean(reduce_precentage, Rx_Medd_2m)
Rx_Medd_3m_new <- transform_mean(reduce_precentage, Rx_Medd_3m)
Rx_Medd_4m_new <- transform_mean(reduce_precentage, Rx_Medd_4m)
Rx_Medd_5m_new <- transform_mean(reduce_precentage, Rx_Medd_5m)
Rx_Medd_6m_new <- transform_mean(reduce_precentage, Rx_Medd_6m)
Rx_Medd_7m_new <- transform_mean(reduce_precentage, Rx_Medd_7m)
Rx_Medd_8m_new <- transform_mean(reduce_precentage, Rx_Medd_8m)
Rx_Medd_9m_new <- transform_mean(reduce_precentage, Rx_Medd_9m)

#combine the new variables with the original variables to create a new dataset
simulated_data <- cbind(Ux_OP_PoC_ED, Ux_OP_PoC_UC, MOPROB_AllAppt, Ux_OP_PoC_TobaccoCessation,  TotalNumOfAppt, Ux_OP_MHOC_HBPC, Rx_OpioidForPain_0to182_D, Rx_OpioidForPain_182to365_D,  Rx_Insomnia_0to365_D, Paton_Rx_Insomnia, Dx_SAE_Acetaminophen_0to1_Y, Dx_SAE_Falls_0to1_Y, Dx_SAE_OtherAccident_0to1_Y, Dx_SAE_OtherDrug_0to1_Y, Dx_SAE_Sedative_0to1_Y, Dx_SAE_Vehicle_0to1_Y, Dx_SedativeUD_0to1_Y, Dx_SUD_CatchAll_RV2_0to1_Y, Dx_AlcoholUD_0to1_Y, Dx_Amphetamine_OtherStim_UD_0to1, Rx_Medd, Rx_Medd_10m, Rx_Medd_11m, Rx_Medd_1m, Rx_Medd_2m, Rx_Medd_3m, Rx_Medd_4m, Rx_Medd_5m, Rx_Medd_6m, Rx_Medd_7m, Rx_Medd_8m,Rx_Medd_9m, agegroup_ls30, agegroup_ge30, agegroup_ge40, agegroup_ge50, agegroup_ge60, agegroup_ge70, sex_female, race_White,  Dx_UDS1_12m, Dx_UDS2_12m, Dx_UDS3_12m,  EH_Psychoses, Dx_MDD_12m, Ux_RehMedPain, Dx_Pain_Urogenital_0to1_Y, Dx_Pain_Neuropathy_0to1_Y, Dx_Pain_Back, Dx_Pain_Neck_0to1_Y, Dx_Pain_Fibromyalgia_0to1_Y, Dx_Endometriosis_0to1_Y, Dx_PTSD_1to2_Y,Dx_Homeless_1to2_Y, Rx_Anxiolytic_0to365_D,  Rx_OAT_methadone_12m, Rx_OAT_buprenorphine_12m, Dx_OpioidOverdose_0to1_Y)

#store the dataset
simulated_data <- as.data.frame(simulated_data)
simulated_data <- data.table(simulated_data)
write.csv(simulated_data, "/Users/luciachen/Desktop/simulated_data/simulated_data_big_sample_reduced_mean.csv")

```